---
title: "R Notebook"
output:
  word_document: default
  html_document:
    df_print: paged
---

## Background

The dataset contains transactions made by European cardholders via credit cards in September 2013, and it includes transactions that occurred over a two-day period, with a total of 284,807 transactions, of which 492 were fraudulent. Also, it provides only the PCA-processed feature values, which are 28 in total. However, it also provides, both time and amount features. Finally, the feature 'Class' is the response variable, which takes the value of 1 in case of fraud and 0 otherwise.\

## Objectives

-   Do Exploratory Data Analysis on the data

-   Conduct cluster analysis

## EDA

First, we have to import packages and the data file we need.

```{r}
library(tidyverse) # data manipulation 
library(cluster) # clustering algorithms 
library(factoextra) # clustering algorithms & visualization
library(caret) # streamline model training process
library(NbClust)
library(corrplot)
library(gridExtra)
library(knitr)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(ggfortify)
dataset = read.csv("creditcard.csv")
```

Before we start, let's take a look at the data.

```{r}
str(dataset)
summary(dataset)
```

We can notice that V1 \~ V28 show nothing special for our analysis, but Class may be a special factor.

## Data preprocessing

```{r}
# scale amount and time
scale_amount = scale(dataset$Amount)
scale_time = scale(dataset$Time)

data_todo = dataset[-c(1,30,31)] %>% 
  mutate(scale_time)%>% 
  mutate(scale_amount) %>%
  mutate(dataset$Class)
```

## Cluster analysis

```{r}
wssplot <- function(data, nc=15, seed=2345){
  wss <- (nrow(data)-1)*sum(apply(data,2,var))
  for (i in 2:nc){
    set.seed(seed)
    wss[i] <- sum(kmeans(data, centers=i)$withinss)}
  plot(1:nc, wss, type="b", xlab="Number of Clusters",
       ylab="Within groups sum of squares")
  wss
}
wssplot(dataset)
```

We use Elbow method to find the k value. It shows above that the result of k is 2. We found that two clusters are the best clusters for this sampled dataset. We then use these two clusters to do k-means calculations.

```{r}
kmean <- kmeans(data_todo, 12)
kmean$centers
autoplot(kmean, data_todo,frame = TRUE)

```

But this result does not represent well the fraudulent and non-fraudulent cases. Cluster 1 represents non-fraudulent cases and Cluster 2 represents fraudulent cases. But they have overlapping parts which are not easily described.
